Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-10-24 manuscript.qmd
add 3D state dist back in

The matrix on the left side of @fig-model-algebra-equation shows the time evolution of the probability distribution over the ipsilateral involvement state, starting from the healthy state $[0,0,0]$ with probability one. The two matrices on the right side show the time evolution of the probability distribution over the contralateral involvement state, where we need to distinguish the two cases with and without midline extension. At time step $t=0$, the contralateral neck starts from the healthy state $[0,0,0]$ without midline extension. The matrix in the center shows the time prior for late T-category. The result of the matrix multiplication is the joint probability distribution over midline extension and ipsi- and contralateral lymph node involvement, $P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon=\texttt{True} \right)$ shown in figure XXX. The joint distribution is shown as two separate heatmaps for the two midline extension states. The most likely state is a lateralized tumor with involvement of ipsilateral level II and no contralateral involvement, which has a probability of approximately 25%. The second most probable state is a lateralized tumor with ipsilateral involvement of levels II and III and no contralateral involvement. The most likely state with contralateral involvement are tumors with midline extension with involvement of contralateral level II and ipsilateral level II and III.

@@

>>>>

@@

The matrix on the left side of @fig-model-algebra-equation shows the time evolution of the probability distribution over the ipsilateral involvement state, starting from the healthy state $[0,0,0]$ with probability one. The two matrices on the right side show the time evolution of the probability distribution over the contralateral involvement state, where we need to distinguish the two cases with and without midline extension. At time step $t=0$, the contralateral neck starts from the healthy state $[0,0,0]$ without midline extension. The matrix in the center shows the time prior for late T-category. The result of the matrix multiplication is the joint probability distribution over midline extension and ipsi- and contralateral lymph node involvement, $P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon=\texttt{True} \right)$ shown in figure @fig-model-state-dist. The joint distribution is shown as two separate heatmaps for the two midline extension states. The most likely state is a lateralized tumor with involvement of ipsilateral level II and no contralateral involvement, which has a probability of approximately 25%. The second most probable state is a lateralized tumor with ipsilateral involvement of levels II and III and no contralateral involvement. The most likely state with contralateral involvement are tumors with midline extension with involvement of contralateral level II and ipsilateral level II and III.

![The full distribution $P \left( \mathbf{X}^\text{i}, \mathbf{X}^\text{c}, \epsilon=\text{False} \right)$ over the ipsi- and contralateral states, as well as midline extension as computed by a model considering the LNLs II, III, and IV. Since this is a three-dimensional object with one dimension being the binary random variable representing midline extension $\epsilon$, we show this distribution as separate two-dimensional heatmaps. These two matrices are the result of @eq-bilateral-marginal-algebra, which we visualized as well in @fig-model-algebra-equation.](figures/fig_model_state_dist.svg){#fig-model-state-dist}
